---
title: "LAB 1"
author: "DT"
date: "2024-09-18"
output: html_document
---

1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called 'Class 1 Survey Fall 2024_di.csv'), calling the R dataframe that you create *C1survey*.


```{r}
#Inputting the survey data into the R dataframe 

url <- "https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv"
C1Survey <- read.csv(url)
```



2. Determine: the number of observations (i.e. the number of people who filled out the survey) and the number of variables in the dataframe.    

There are 29 observations and 27 variables
```


3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about 
b. Display that they are renamed.


```{r}
# a. Rename variables to be descriptive and shorter.
names(C1survey)[1:27]<- c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

# b. Display renamed variables 
colnames(C1survey)
```



```

4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset


```{r}
#a. Determine the class of variable in the data set.

sapply(C1survey, class)

str(C1survey) 
table(sapply(C1survey, class))
summary(C1survey)
```


```


5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values. If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}


#a 
C1survey$bday
C1survey$bmonth

#b 
## Here we are recoding the birthday days. So the number of the day is the recorded response. 
class(C1survey$bday) 
C1survey$bday<-ifelse(C1survey$bday == "May 21-report 21", "21", C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday == "March 31st", "31", C1survey$bday)
C1survey$bday 

#This code converts the character variables to numeric variables. 
#The table function creates a table for the values that are NA. 
C1survey$bday<-as.numeric(C1survey$bday) 
table(C1survey$bday, useNA = "always") 

#Then we do the same thing for birthday month. 
#First, we check the variables in bmonth to see the class they fall into.
class(C1survey$bmonth)

#We rocode the variables in bmonth to represent the numeric value of the associated month. 
C1survey$bmonth[C1survey$bmonth=="May 21-report 5"]<-"5"
C1survey$bmonth[C1survey$bmonth=="February"]<-"2"
C1survey$bmonth[C1survey$bmonth=="September"]<-"9"
C1survey$bmonth[C1survey$bmonth=="July"]<-"7"
C1survey$bmonth[C1survey$bmonth=="March"]<-"3"
C1survey$bmonth 

#Now we convert bmonth to numeric values and view the numeric values.
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth 

#c. 
median(C1survey$bday, na.rm=TRUE) 
median(C1survey$bmonth, na.rm=TRUE)
```


```


```6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 


```{r}

#a. Create a new variable called *bseason* that gives the season
C1survey <- C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(C1survey$bmonth, C1survey$bseason) 

C1survey$bseason_f<-factor(C1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))


#b. Print a table of Seasons in the columns and bmonths in the rows
table(C1survey$bmonth, C1survey$bseason)
table(C1survey$bmonth, C1survey$bseason_f)

#c. Sum the columns using the addmargins function
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason_f, useNA = "always"), 1)
tab

# In our class, 8 people were born in the Winter. 9 people were born in the Spring. 8 people were born in the Summer and 4 people were born in the Fall.

```




7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).


```{r}
install.packages("dplyr")
library(dplyr)

#Here we created a data frame (month_to_season) the corresponds months to the season that they are in (example: January and Winter)
month_to_season <- data.frame(
  Month = c("January", "February", "March", "April", "May", "June", 
            "July", "August", "September", "October", "November", "December"),
  Season = c("Winter", "Winter", "Spring", "Spring", "Spring", "Summer",
             "Summer", "Summer", "Fall", "Fall", "Fall", "Winter")
)

#Next we changed the survey set to contain the new data frame month to season. Then we applied the data frame month to season to identify the season in which the respondents favorite month resides.

C1survey_with_season <- C1survey %>%
  left_join(month_to_season, by = c('fav_month' = "Month"))

#Next we created a new column called MonthMatchesSeason to see if the respondent's favorite season corresponds to the season in which their favorite month is.
C1survey_with_season <- C1survey_with_season %>%
  mutate(MonthMatchesSeason = `fav_season` == Season)

#Next we want to print out the overall percent matches from all 29 responses.
summary_results <- C1survey_with_season %>%
  summarise(PercentageMatch = mean(MonthMatchesSeason, na.rm = TRUE) * 100)

#We want to then print the result which is the percent of all respondent's whose favorite season matches the season of their favorite month.
print(summary_results)

#Finally, we want to stratify the results by season.
season_match_percent <- C1survey_with_season %>%
  group_by(`fav_season`) %>%
  summarise(
    Total = n(), #This is the # of respondents who responded for a specific season
    Matches = sum(MonthMatchesSeason, na.rm = TRUE), #This is the total matches within a specific season whose favorite month resides in their favorite season.
    PercentMatch = (Matches / Total) * 100
  )

#This is the output of percent matches by season.
season_match_percent
```


